---
title: "Map Geospatial and Demographic Data"
author: "Veronica Ikeshoji-Orlati"
date: "1/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r}
# Import libraries
library(tidyverse)
library(leaflet)
library(sf)
library(raster)
library(mapview)
```

## The 2000 and 2010 Census / ACS Tracts

There is a slight discrepancy between 2000 and 2010 census tracts. As a result, I have downloaded both shapefiles from [the US Census Bureau's Mapping Files service](https://www.census.gov/geographies/mapping-files.html) and import them here.
```{r}
# Import the tract shapefile for 2000 and 2010
tract_2000 <- st_read("data/2000_census_tract_map/tl_2010_47037_tract00.shp")
tract_2010 <- st_read("data/2010_census_tract_map/tl_2010_47037_tract10.shp")

# Fix the CRS to bring it into alignment with the other geospatial datasets
tract_2000 <- st_transform(tract_2000, CRS("+proj=longlat +datum=WGS84 +no_defs"))
tract_2010 <- st_transform(tract_2010, CRS("+proj=longlat +datum=WGS84 +no_defs"))

# Check the CRS for both
crs(tract_2000)
crs(tract_2010)

# Map the 2000 census tracts for Davidson County
nash_2000 <- leaflet() %>%
  leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng=-86.7410562, lat=36.1682906, zoom = 11) %>% 
  addPolygons(data=tract_2000,
              weight=2,
              col = 'red',
              popup = tract_2000$TRACTCE00)

nash_2000

# Map the 2010 census tracts for Davidson County
nash_2010 <- leaflet() %>%
  leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng=-86.7410562, lat=36.1682906, zoom = 11) %>% 
  addPolygons(data=tract_2010,
              weight=2,
              col = 'blue',
              popup = tract_2010$GEOID10)

nash_2010

# Overlap the maps
nash_census_map <- leaflet() %>%
  leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng=-86.7410562, lat=36.1682906, zoom = 11) %>% 
  addPolygons(data=tract_2000,
              weight=4,
              col = 'red',
              popup = tract_2000$TRACTCE00) %>% 
    addPolygons(data=tract_2010,
              weight=2,
              col = 'blue',
              popup = tract_2010$GEOID10)
nash_census_map

# NB: In 2000, the census tract ID is CTIDFP00. In 2010, the census tract ID is GEOID. It is unclear why this had to change, nor why it takes a dive into the technical documentation to determine that they are the same thing.

# Find the tractIDs that are only in one dataframe or the other
# Try it spatially first
test_sdf <- st_join(
  tract_2000,
  tract_2010,
  join = st_contains,
  suffix = c(".x", ".y"),
  left = TRUE,
  largest = FALSE
)
test_sdf_map <- leaflet() %>%
  leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng=-86.7410562, lat=36.1682906, zoom = 11) %>% 
  addPolygons(data=test_sdf,
              weight=2,
              col = 'red')

test_sdf_map
# Write out the spatial dfs to CSVs
write_csv(tract_2000, "data/tract_2000.csv")
write_csv(tract_2010, "data/tract_2010.csv")

# Read them back in as non-spatial dfs
tract_2000_df <- read_csv("data/tract_2000.csv")
tract_2010_df <- read_csv("data/tract_2010.csv")

# Try again using the dplyr::anti_join() function, only get the tracts that are in the 2000 and 2010 census
unmatched_2000_tracts <- anti_join(tract_2000_df, tract_2010_df, by = c("TRACTCE00" = "TRACTCE10"))
unmatched_2010_tracts <- anti_join(tract_2010_df, tract_2000_df, by = c("TRACTCE10" = "TRACTCE00"))

# Write them out to CSVs
write_csv(unmatched_2000_tracts, "data/unmatched_2000_tracts.csv")
write_csv(unmatched_2010_tracts, "data/unmatched_2010_tracts.csv")


# Based on the overlap, normalize to the 2000 census tracts by creating a new GEOID column for the 2010 data
```

## The Tornado Dataset

I downloaded the tornado dataset from the NOAA; here is the [documentation](https://www.spc.noaa.gov/wcm/data/SPC_severe_database_description.pdf) for the dataset, which I acquired [here](https://oasishub.co/dataset/usa-tornado-historical-tracks-noaa/resource/b2a11100-eac5-4d10-869a-87ba064ede2d?inner_span=True).
```{r}
# Read in the tornadoes dataset
tornadoes <- read.csv("data/TN_tornadoes_1950-2018.csv")

# Filter it down to only include tornadoes since 1990
tornadoes <- tornadoes %>% 
  filter(yr %in% (1990:2020))

# Create a uid for each tornado
tornadoes$uid <- paste(tornadoes$yr,tornadoes$om,sep = "_")  

# Pare down the dataset to only the requisite variables
tornadoes <- tornadoes %>% dplyr::select(uid, date, slat, slon, elat, elon, mag)

# For tornadoes with no end lat/long, replace with start lat/long (these are the touchdowns that don't travel)
tornadoes$elat[tornadoes$elat=="0"] <- tornadoes$slat[tornadoes$elat == "0"]
tornadoes$elon[tornadoes$elon=="0"] <- tornadoes$slon[tornadoes$elon == "0"]

# Create a new tornado label
# TODO: tornadoes$popup <- paste(date, (mag), sep = ", magnitude ")

# Write out the cleaned dataframe to a csv to import into the app
write_csv(tornadoes, "data/tornadoes_for_mapping.csv")

# Create a base map for the tornadoes dataset
tornado_map <- leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>% 
  setView(lng=-86.7410562, lat=36.1682906, zoom = 9.5)

# Check the map
tornado_map

# Write a for loop to create a new line for each tornado path
for (i in 1:nrow(tornadoes)) {
  tornado_map <-
  tornado_map %>%
  addPolylines(lat=c(tornadoes[i,]$slat,tornadoes[i,]$elat),
               lng=c(tornadoes[i,]$slon,tornadoes[i,]$elon),
               popup = paste("Date:", tornadoes[i,]$date, "<br>",
                           "Magnitude:", tornadoes[i,]$mag))
}
tornado_map
```

## Read in the HOLC Dataset and Map It

The Home Owners' Loan Corporation was a government agency that, between 1935 and 1940, assigned 'mortgage security' grades to neighborhoods in metropolitan areas throughout the US. The [Mapping Inequality: Redlining in New Deal America](https://dsl.richmond.edu/panorama/redlining/#loc=5/39.1/-94.58) project at the University of Richmond has georectified these historic maps and made them available for download on the project website; their shapefiles for Nashville are utilized here.

``` {r}
# Read in the HOLC dataset
holc <- st_read("data/HOLC_BNA_shape_files/cartodb-query.shp")

# Transform it and make sure it's using the right projection
holc <- st_transform(holc, CRS("+proj=longlat +datum=WGS84"))

# Create a color-coding palette based on the HOLC data
holc_pal <- colorFactor(
  palette = c('#4ca64c', '#4d4dff', '#ffff7f', '#ff1a1a'),
  domain = holc$holc_grade,
  alpha = TRUE # This ensures that the transparency arguments are utilized in the addPolygons function
)

# Map it
holc_map <- leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = holc,
              fillColor = ~holc_pal(holc_grade),
              #color = "#FF7F50", # you need to use hex colors
              fillOpacity = 0.5, 
              weight = 0, 
              smoothFactor = 0.2,
              popup = holc$holc_grade)

# Take a look at the map
holc_map
```

## Map it all together

``` {r}
nash_map <- leaflet() %>%
  leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng=-86.7410562, lat=36.1682906, zoom = 11) %>% 
  addPolygons(data=tract_2000,
              weight=4,
              col = 'red',
              popup = tract_2000$TRACTCE00) %>% 
    addPolygons(data=tract_2010,
              weight=2,
              col = 'blue',
              popup = tract_2010$GEOID10) %>% 
    addPolygons(data = holc,
              fillColor = ~holc_pal(holc_grade),
              #color = "#FF7F50", # you need to use hex colors
              fillOpacity = 0.5, 
              weight = 0, 
              smoothFactor = 0.2,
              popup = holc$holc_grade)

for (i in 1:nrow(tornadoes)) 
  nash_map <-
  nash_map %>%
  addPolylines(lat=c(tornadoes[i,]$slat,tornadoes[i,]$elat),
               lng=c(tornadoes[i,]$slon,tornadoes[i,]$elon),
               popup = tornadoes[i,]$date)
nash_map
```

